---
title: An R Markdown document converted from "copy_of_VectorsIn_R.ipynb"
output: html_document
---

# **VECTORS in R**

## 1. Creating

- You use **c()** to create a vector
- Every vector element should share the **same data type**

```{r}
# creating 5 vector

namesR=c("Qing", "Françoise", "Raúl", "Bjork","Marie")
agesR=c(32,33,28,30,29)
countryR=c("China", "Senegal", "España", "Norway","Korea")
educationR=c("Bach", "Bach", "Master", "PhD","PhD")
```

```{r}
namesR
```

## 2. Accessing

The vector is a **container**. Accessing means **seeing** *one* or *several* values, seeing a value changes nothing in the vector.

```{r}
# one value
agesR[1]
```

```{r}
# several
agesR[c(1,3)]
```

### Slicing

You use the **:** to request some values:

```{r}
# consecutive
agesR[2:4]
# agesR[seq(2,4)]
```

```{r}
# non consecutive - but with a pattern
agesR[seq(1,4,2)]
```

** Not sure what the sequence is or the pattern **

## Replacing

Here the list will change some contents, but the **amount** of elements stays the same.

### by Position

We tell the location, or **index**, of the element to change:

```{r}
# remember: countryR=c("China", "Senegal", "España", "Norway","Korea")

countryR[3]="Spain"

# list changed:
countryR
```

### By Value

We find a match to a current value, and update it with different one:

```{r}
countryR[countryR=='China']="PR China"
countryR
```

## Deleting

Here the vector will shrink.

Notice we are rewriting the same object:

### By Position

```{r}
# if we have countryR=c('PR China','Senegal','España','Norway','Korea')

#this is a VIEW
countryR[-5] # bye Korea
```

```{r}
# Korea still there?
countryR
```

```{r}
# this is rewriting
countryR=countryR[-5]

# vector object rewritten
countryR
```

We can delete more than one position:

```{r}
vectorR=c(1,2,3,4,5,6,100,10000) #new vector created

vectorR=vectorR[-c(2:4)]

#now:
vectorR
```

You can **filter** instead of deleting:

```{r}
# non consecutive

vectorR=c(11,22,33,44,55,66) #new vector created

new_vectorR=vectorR[-c(2,4,6)] #another new vector created
#now
new_vectorR
```

```{r}
#remains intact
vectorR
```

### by Value

```{r}
#one value

vectorR=c(1,3,45,3,3) #new vector created

vectorR=vectorR[vectorR!=3]

#then
vectorR
```

Notice the use of **%in%**:


```{r}
#several values

vectorR=c(1,3,45,3,3)#new vector created

vectorR=vectorR[!vectorR%in%c(1,45)]
vectorR
```

## Inserting values

Vector will become larger:

### at the End

```{r}
vectorR=c(1,2,3,4) #new vector created

vectorR=c(vectorR,5)

# now
vectorR
```

### in a particular location

```{r}
vectorR=c(1,2,3,4) #new vector created

vectorR=append(vectorR,10,after=1)

# now
vectorR
```

## Concatenating

```{r}
vectorR1=c(1,2,3,4) #new vector created
vectorR2=c('a','b') #another new vector created
```

This creates a new vector:

```{r}
#another new vector created
vectorR12=c(vectorR1,vectorR2)

# now
vectorR12
```

This make one vector grow larger:

```{r}
vectorR1=c(vectorR1,vectorR2)

# now
vectorR1
```

Remember that vectors support just one data type, and coercion will activate when different data types are detected.

```{r}
vectorR1=c(1,2,3,4,"o")
vectorR1
```

